---
title: "WUP_04"
author: "Jarrod Harris"
date: "Summer 2025"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
mydata <- read.csv(file="GSScsv.csv")
```
# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->



# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(GSScsv,aes(x=WORDSUM,y=DEGREE))

basicCN + geom_boxplot()
```


In the plot the LT high and high school is under the jr college and bachelor schools. This plot shows 2 boxes are beneath the others and that they probably have a statistical difference 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=mydata)
```

In the data it shows that the LT high school and the high school Q3 is lower than the graduate schools Q1. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=mydata)
summary.aov(model1)
```

the P value data is a small number of 0.000236

```{r}
par(mar = c(3, 17, 2, 2))
plot(TukeyHSD(model1, conf.level =0.98), las=2)
```

From viewing the graph i can tell its a statistical difference between the high school graduates and college graduates, as well as the difference in the LT high school and graduate. Also the dotted line thats shown above 0 is the only thing that doesn't interfere with the LT graduates or the other graduates. 




# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
